---
title: "Getting started with accelerometer"
author: "Tuomo Nieminen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

To access the accelerometer package, use:

```{r}
library(accelerometer)
```


## Reading .gt3x data into R

The accelerometer package includes two sample .gt3x files which I'll use to demonstrate reading the data. First we need the path to a single gt3x file.

```{r}
gt3xfile <- gt3x_datapath(1)
```


### Method 1: Temporary unzip and read

readGT3X() can take as input a path to a single .gt3x file and will then read activity samples as an R matrix.

```{r}
X <- readGT3X(gt3xfile)
```

```{r}
head(X)
```


### Method 2: permanent unzip, then read

.gt3x files are actually zip archives which contain two files: log.bin and info.txt. log.bin is a binary file that contains the actual samples. It usually make sense to store the data as *unzipped* folders containing these two files, because otherwise the readGT3X() function will have to unzip each .gt3x archive every time you need to access the data. 

The unzipGT3X() function unzips all .gt3x files in a given directory. By default, the contents of a .gt3x file named "subject001.gt3x" are extracted to a folder named "subject001". unzipGT3X() returns a vector of paths to each of the unzipped gt3x folders. The location argument can be used to choose where to locate those folders.

```{r}
datadir <- gt3x_datapath()
gt3xfolders <- unzipGT3X(datadir, location = tempdir())
gt3xfolder <- gt3xfolders[1]
```

The readGT3X function also accepts a path to an uzipped gt3x folder. It is a bit faster if the unzip step has already been performed.

```{r}
X <- readGT3X(gt3xfolder)
```

```{r}
head(X)
```

## Activity data matrix

Internally, the data matrix returned by readGT3X is a bit smarter than it looks, as it knows all the (relative) timestamps of the observations.  

```{r}
str(X)
```
